---
title: "Info on eat Packages"
---

see table down below and click on the links for more information on a package.

## internal IQB R packages from the 'eat'-family

In order to standardise and simplify the evaluation of the VERA pilot studies and the IQB educational trend, a number of R packages with the prefix eat ("educational assessment tools") are continuously (further) developed. Currently, the following packages are maintained at the IQB:

```{r, include=FALSE}
#| Paketname       | Ansprechpartner:in                                  | mehr Infos                          |
#|-------------------|----------------------------------|-------------------|
#| **eatPrep**     | Karoline Sachse                                     | [more Info on eatPrep](eatPrep.qmd) |
#| **eatModel**    | Sebastian Weirich                                   | [more Info on eatModel](eatModel.qmd) |
#| **eatRep**      | Sebastian Weirich, Benjamin Becker                  | [more Info on eatRep](eatRep.qmd) |
#| **eatGADS**     | Benjamin Becker                                     | [more Info on eatGADS](eatGADS.qmd) |
#| **eatTools**    | Sebastian Weirich, Benjamin Becker, Karoline Sachse | [more Info on eatTools](eatTools.qmd) |
#| **eatAnalysis** | Benjamin Becker, Sebastian Weirich, Karoline Sachse | [more Info on eatAnalysis](eatAnalysis.qmd) |
#| **eatATA**      | Benjamin Becker                                     | [more Info on eatATA](eatATA.qmd) |
#| **eatFDZ**      | Benjamin Becker                                     | [more Info on eatFDZ](eatFDZ.qmd) |
#| **eatCodebook** | Benjamin Becker                                     | [more Info on eatCodebook](eatCodebook.qmd) |
#| **eatRecode**   | Benjamin Becker, Nicklas Hafiz                      | [more Info on eatRecode](eatRecode.qmd) |
#| **eatPlot**     | Nicklas Hafiz, Philipp Franikowski                  | [more Info on eatPlot](eatPlot.qmd) |
```



```{r, echo=FALSE}
#Paketname <- c("eatPrep", "eatModel", "eatRep", "eatGADS", "eatTools", 
#               "eatAnalysis", "eatATA", "eatFDZ", "eatCodebook", "eatRecode", 
#               "eatPlot")
links <- ""
links[1] <- "[eatPrep](eatPrep.qmd)"
links[2] <- "[eatModel](eatModal.qmd)"
links[3] <- "[eatRep](eatRep.qmd)"
links[4] <- "[eatGADS](eatGADS.qmd)"
links[5] <- "[eatTools](eatTools.qmd)"
links[6] <- "[eatAnalysis](eatAnalysis.qmd)"
links[7] <- "[eatATA](eatATA.qmd)"
links[8] <- "[eatFDZ](eatFDZ.qmd)"
links[9] <- "[eatCodebook](eatCodebook.qmd)"
links[10]<- "[eatRecode](eatRecode.qmd)"
links[11]<- "[eatPlot](eatPlot.qmd)"

desc <- ""
desc[1] <- "Preparation of data sets."
desc[2] <- "Interface for ConQuest software."
desc[3] <- "calculates characteristics for data with clustered multi-level structures with imputed data."
desc[4] <- "Import and data preparation of SPSS data sets in R."
desc[5] <- "Various help functions that are also required by the packages 'eatPrep', 'eatModel', 'eatGADS' and 'eatRep', among others."
desc[6] <- "Various useful help functions, such as saving Excel files, saving analysis results from lm4 and simulating IRT responses."
desc[7] <- "Automated block occupation/automated test booklet creation."
desc[8] <- "Automated anonymisation of data sets, matching of pdf documents (e.g. scale manuals) and data sets."
desc[9] <- "Automated creation of scale manuals."
desc[10]<- "Create and apply recoding databases."
desc[11]<- "Creating plots from the eatRep output (main use: Bilduntstrend)."

an <- ""
an[1] <- "Karoline Sachse"
an[2] <- "Sebastian Weirich"
an[3] <- "Sebastian Weirich, Benjamin Becker"
an[4] <- "Benjamin Becker"
an[5] <- "Sebastian Weirich, Benjamin Becker, Karoline Sachse"
an[6] <- "Benjamin Becker, Sebastian Weirich, Karoline Sachse"
an[7] <- "Benjamin Becker"
an[8] <- "Benjamin Becker"
an[9] <- "Benjamin Becker"
an[10]<- "Benjamin Becker, Nicklas Hafiz"
an[11]<- "Nicklas Hafiz, Philipp Franikowski"

df <- data.frame(links, desc, an)

# find out how to print nicer
df_layout <- df
colnames(df_layout) <- c("package name", "brief description", "contact")

#df_layout
knitr::kable(df_layout)

```


## Installing R packages

Detailed instructions for installing the R packages can be found here: [installing R packages](http://iqbstaff.pbworks.com/w/page/138429033/R-Pakete%20Installieren)

## Compatibility

Most packages of the "eat" family are not executable in isolation, which means that a package is usually dependent on another package in order to be executable. This means:

- "eatTools" is a package with help functions required by other packages. "eatTools" does not depend on other "eat" packages.
- eatPrep" requires "eatTools".
- eatRep" needs "eatTools" and "eatGADS".
- eatModel" requires "eatTools" and "eatRep" (and thus also "eatGADS"). The package optionally accesses the computer programme Conquest or the R package "TAM" for parameter estimation. For Conquest, the command line executable is sufficient, e.g. "console_Feb2007.exe"; "TAM" is installed when the package is loaded. "TAM" and Conquest are based on the same statistical measurement model, and overlap considerably in their range of functions.
- "eatGADS" requires "eatTools" and "eatDB".

The interdependence also extends to different versions of the packages. For example, old versions of "eatRep" are not always compatible with new versions of "eatTools" (and vice versa). The current package versions (see table above) should be compatible with each other. Usually (but not always) version conflicts are indicated by an error message. It is recommended to always have the latest package versions installed.

Since individual functions have also changed substantially in the course of the package development, it may no longer be possible to replicate past analyses with new package versions and the identical script (e.g. from the Ländervergleich 2011). In this case, the old script would either have to be adapted or the package versions used at that time would have to be restored.
