---
title: "Reproducable manuscripts with Quarto"
description: "R-SIG 15.07.2024"
author: 
  - name: Nicklas Hafiz
  - affiliation: PhD student at the IQB, Methods team
categories: [R, quarto]
date: 07-15-2024
bibliography: references.bib
bibliographystyle: apa
---

# Introduction

-   [Quarto](https://quarto.org/) is an open-source scientific and technical publishing system.
-   It allows you to combine code and markdown text to produce reproducable manuscripts that can automatically incorporate and evaluate your code when rendering. No mor copy-pasting from tables. No more screenshoting plots. No more trying to get the references formatted correctly. And much more!
-   One great thing about quarto: because you write your stuff in Markdown language (it's not hard), you are free to choose into which format to export, and can change that easily. I could for example export my manuscript as PDF, and with just changing one word, I can export into word, HTML ...
-   You can also build slides and even websites (like the one you are currently on) with quarto.
-   Great stuff!

# Workflow

-   RStudio project
-   Optional (but highly recommended): GitHub
-   Optional (but highly recommended): [renv](https://iqb-research.github.io/IQB-Methods/posts/r_sig/23_11_06_renv/)
-   `File - New File - Quarto Document` manuscript (best is to split up the manuscript into chapters)
-   R-Code either seperatly or within your manuscript
-   Personal preference: Splitting up pretty much, makes it easier to keep an overview.

# Components

## YAML header

Begin on top of document by writing `---`, and also end with `---`. It contains global options for your document:

```         
---
title: "Reproducable manuscripts with Quarto"
description: "R-SIG 15.07.2024"
author: 
  - name: Nicklas Hafiz
  - affiliation: PhD student at the IQB, Methods team
categories: [R, quarto]
date: 07-15-2024
---
```

An overview of possible YAML-fields can be found [here](https://quarto.org/docs/reference/formats/pdf.html).

## Markdown text

Write your text in markdown language. Some expressions:

-   `#` for headers. Add as many `#` as you like for subheaders.
- `**bold**, *italic*, `code``
- Lists: `-` for bullet points, `1.` for numbered lists (beware: the line above the list has to be empty)
- Linebreaks: Two spaces at the end of a line. 
- Links: `[text](url)`

## Code chunks


```{r}
#| echo: true

library(palmerpenguins)


```

Chunk options: 




## Tables
Tables can be a bit tricky in quarto. The great thing is: once you have styled them, they get updated automatically if some data changes.  
For some simple tables, the normal markdown syntax might be enough, see [here](https://quarto.org/docs/authoring/tables.html). I'd like to point out `knitr::kable()`, which allows some extra styling. 
For more complex tables, I've found the [rempsyc](https://rempsyc.remi-theriault.com/articles/table) package which helps in building APA tables, but there are other options like [flextable](https://davidgohel.github.io/flextable/) or [gt](https://gt.rstudio.com/)  as well. 



## Citations
Citations are saved in `.bib` files. The `.bib` format can be chosen on almost every website, often next to APA and others. It looks like this: 

```
@article{allport1936trait,
  title={Trait-names: A psycho-lexical study.},
  author={Allport, Gordon W and Odbert, Henry S},
  journal={Psychological monographs},
  volume={47},
  number={1},
  pages={i},
  year={1936},
  publisher={Psychological Review Company}
}

@book{darwin1859,
  added-at = {2008-05-27T04:02:47.000+0200},
  address = {London},
  author = {Darwin, Charles},
  biburl = {https://www.bibsonomy.org/bibtex/2d70d713c717fb28384fb073c9f6dfbc2/neilernst},
  citeulike-article-id = {2376343},
  interhash = {c738acbb887362be5b0e6abc51be42d3},
  intrahash = {d70d713c717fb28384fb073c9f6dfbc2},
  keywords = {evolution},
  note = { or the Preservation of Favored Races in the Struggle for Life},
  priority = {2},
  publisher = {Murray},
  timestamp = {2008-05-27T04:02:47.000+0200},
  title = {On the Origin of Species by Means of Natural Selection},
  year = 1859
}

```
The great thing is: from this format, you can generate all kinds of different citation styles. Just put `bibliographystyle:` with the wanted style into your YAML-header: `bibliographystyle: apa`

Of course, you should still check if the fields are filled in correctly.
So if you have created a `references.bib` file in your project directory, you can include it in your quarto document by adding `bibliography: references.bib` to your YAML header.  
To cite a reference in you text, you can use an @ in front of the tag like `@darwin1859`: @darwin1859.



# Templates & APA

## Own Templates
You can always create your [own template](https://quarto.org/docs/journals/templates.html) Luckily, some people have done a lot of the work for us, and therefore we can use templates provided for [specific journals](https://github.com/quarto-journals/). Oooor more generally, APA templates: 

## Papaja
There are some markdown templates that format your text in APA style. Most famously, the [`papaja`](https://github.com/crsh/papaja) package lets you write APA conform manuscripts. Sadly only in R Markdown, not in Quarto.   

## apaquarto
Alternatively, I've found (but not yet tested in a whole project) the [apaquarto](https://wjschne.github.io/apaquarto/) extension. You can install via:

```{r}
#| eval: false
#| echo: true

setwd("home/my_project") # Make sure the folder is empty

quarto::quarto_use_template("wjschne/apaquarto")
```

This will create the necessary files in your folder. Update the `.qmd` file with the same name as your folder. I'd suggest to also create a RStudio-project there, and maybe structure your files into multiple folders (data, R-Scripts ...).  
The template should label your tables and figures correctly, and format the bibliography as well as the whole document. 


# Exercises

0. Create a fresh, empty folder for this exercise.
1. Install the `apaquarto` template. 

::: {.callout-tip collapse="true"}
## Hint

- Make sure you have a current version of quarto installed. If not, download from [here](https://quarto.org/docs/download/) and install. Restart RStudio, it should find the new quarto installation automatically. 
- Make sure to set the working directory to an empty folder before installing: 

```{r}
#| eval: false
#| echo: true

setwd("home/my_project") # Make sure the folder is empty
quarto::quarto_use_template("wjschne/apaquarto")

```


:::

2. Create a RStudio project in your directory and open it. 
3. Open the `.qmd` document and fill it with a "mini"-analysis of the [`athletes`](https://github.com/nickhaf/r_tutorial/blob/main/raw_data/athletes.rds) dataset. So update the YAML and the text body for your own needs. The analysis should contain: 

- at least two chapters
- some sort of table
- some plot
- two own references (a book and an article)

Everything should be in APA7 style.   
Feel free to use summaries and plots from the previous [R-SIG](https://nickhaf.github.io/r_tutorial/qmd/plotting/plotting.html). 



::: {.callout-caution collapse="true"}
## Solution

```{r, message=FALSE}
#| echo: true

library(tidyverse)

athletes <- readRDS(file = here::here(  "raw_data", "athletes.rds"))

best_by_sport <- athletes %>%
  ## Get all gold medalists
  filter(Medal == "Gold") %>%
  ## Group them by sport and region
  group_by(Sport, Region) %>%
  ## count the number of medals each country has per sport category
  count(Medal) %>%
  ## Now only group by sport, so we can extract the maximum medal row by sport, and not by sport and country
  group_by(Sport) %>%
  ## Extract the country with the most medals
  slice(which.max(n))

head(best_by_sport)

```


```{r}
world_coordinates <- readRDS(file = here::here("raw_data", "world_coordinates.rds"))

medal_counts <- athletes %>%
  filter(Medal == "Gold") %>%
  group_by(Region) %>%
  count(Medal)

medal_countries <- world_coordinates %>%
  rename("Region" = region) %>%
  left_join(medal_counts)

p <- ggplot(
  data = medal_countries,
  mapping = aes(x = long, y = lat, group = group)
) +
    geom_polygon(aes(fill = n)) +
    coord_fixed(1.3) +
  ## Use another colour palette:
  scale_fill_distiller(palette = "YlOrBr", direction = 1) +
  ## Plot title
  ggtitle("Olympic Gold Medal winners") +
  ## Legend title:
  guides(fill = guide_legend(title = "Gold medal \n winners")) +
  ## Minimal theme:
  theme_void()

p
```

:::

4. Split up your document into multiple files. This is kind of a preference thing, however I find it easier to keep an overview if I split up my chapters and R code and don't have a huge `.qmd` file containing everything. So try it! Create a folder called `docs` or something similar. 