---
title: "Using Checkmate"
author: "Edna Grewers"
date: 2023-12-01
---

# Intro

With the checkmate package you can test, check and assert all kinds of arguments regarding type, length and much more. You can also write your own assert-functions.

In this Vignette you can get a broad understand how the package works, how the output looks like and some example functions. For more details and a complete overview go to [CRAN](https://cran.r-project.org/web/packages/checkmate/index.html), specifically the [checkmate.pdf](https://cran.r-project.org/web/packages/checkmate/checkmate.pdf) or use the following line of code:

```{r, eval=FALSE}
lsf.str("package:checkmate")
```

## Setup

Install the package from `CRAN` with the following code. Then load it in your library.

```{r, eval=FALSE}
install.packages("checkmate")
```

```{r}
library(checkmate)
```


## Different Outputs

There are three main functions that we use: `test`, `check` and `assert`, which do similar things but produce different outputs. All functions have two ways to write them: `test_numeric()` with an underscore and `testNumeric()` with a capital Letter, but they both work the same. For simplicity we'll just use one option.

To show the differences in outputs we check arguments for numeric input, as an example. See below to see checks for different types or attributes.

**Test**

Test-functions test whether an argument has certain attributes and gives you `TRUE` or `FALSE` output.

```{r}
test_numeric(c(6:1, 4))
test_numeric("hallo") 
```

**Check**

Check-functions check whether an argument has certain attributes and gives you `TRUE` or a `string` containing an error message as an output. In the string you can see what kind of argument you should have given, and what you did wrong.

```{r}
check_numeric(c(6:1, 4))
check_numeric("hallo") 
```

**Assert**

Assert-functions assert whether an argument has certain attributes and throw an `error message` if you it doesn't. When you did everything correct, it doesn't create an output. The `error message` contains the `string` from the `check`-functions.

```{r, error=TRUE}
assert_numeric(c(6:1, 4))
assert_numeric("hallo")
```

If you save an `assert_numeric()` object into a variable x, it will contain the original object that you asserted.

```{r}
x <- assert_numeric(c(6:1, 4))
x
```

## Checking for Type

With `lsf.str()` you can see all functions of the package.

```{r, eval=FALSE}
lsf.str("package:checkmate")
```

You can check for specific types of arguments e.g. `numeric`, `number`, `integer`, `double`, `character`, `string`, `logical`, `flag`, `missing`, or data structure e.g. `list`, `data_frame`, `array` and so on. You can also look for attributes e.g. `true`, `subset`, `named` or `atomic`, and much more.

Depending on what kind of output you want, you choose your function.

```{r}
check_list(list())
check_list(1:9)
```

If you are looking for a specific function you can use:

```{r}
ls("package:checkmate", pattern = "atomic") 
```

### Checking for Type: Data Frames

If you want to check the types of elements of a more complex data structure like a `list` or a `data frame`, you have to look at the arguments of the functions. Both have the argument `types`.

First we create an example data frame with rows and columns named. It has numeric and character elements.

```{r}
df <- data.frame(klein = 1:3, mittel = 4:6, groß = c("7", "8", "9"), row.names = c("A", "B", "C"))
df
```

Now we can check for the types of the elements using the `types` argument. You can look at the whole data frame with `df`, at single columns using `df[1]` or `df[klein]` or at single rows using `df[1,]`. The error message will tell you the first element that has a type you didn't check for.

```{r}
# checking the whole data frame
check_data_frame(df, types = c("numeric", "character"))
check_data_frame(df, types = "numeric")

# checking individual columns
check_data_frame(df[1], types = "numeric")
check_data_frame(df["klein"], types = "numeric")
check_data_frame(df[3], types = "numeric")

# checking for individual rows
df[1,]
check_data_frame(df[1,], types = "numeric")

```

You can also check the type of individual elements with `df[1,1]` or just the content of the columns by loosing the attributes of the data frame with `df$klein` or `df[,1]`. In this example they all have type `integer`. Now `check_data_frame()` doesn't work anymore, because the data is no longer a data frame. You can use the normal checks from above.

```{r}
check_data_frame(df$klein, types = "numeric")

check_integer(df$klein)
check_numeric(df[,1])
check_double(df[1,1])
```

### Checking for Type: Lists

Lists work the same way. We create an example list a, with named elements of different types.

```{r}
a <- list(zahlen = 1:9, mon = month.abb, creator = "IQB")
assert_list(a, types = c("numeric", "character"))
assert_character(a$mon)
```

## Checking for Length

You can check if an argument is `scalar`, but you cannot check for arbitrary lengths this way.

```{r, error=TRUE}
check_scalar(1)
check_scalar(1:5)

ls("package:checkmate", pattern = "length")
```

To check for arbitrary length of an argument, you have to use the `len` argument from the `test`, `check` or `assert` functions.

```{r, error=TRUE}
check_character(month.abb, len = 12)
check_character(month.abb, len = 11)
```

There are some other attributes you can check like this, e.g. `min.len`, `max.len`, `unique` or the length of elements in character vectors using `n.chars`, `min.chars` or `max.chars`.

```{r}
check_character(month.abb, n.chars = 3, min.chars = 2)
check_character(month.abb, n.chars = 2)
check_character(month.abb, n.chars = 3, max.chars = 2)
```

For more info what arguments you can check, type `?check_character` in the console.

```{r, eval=FALSE}
?check_character
```

You can also check the lengths of **lists** or the length of columns/rows of a **data frame** in a similar way by using our example objects from above.

```{r}
assert_list(a, len = 3, min.len = 2, max.len = 3)
assert_data_frame(df, min.cols = 1, max.cols = 3, ncols = 3)
```

## Checking Names via Subset

You cannot check the names of complex objects directly. With the **list** or **data_frame** functions you can only check if the objects is named at all, or if the names are unique via the arguments `names` for lists and `col.names` or \``row.names` for data frames. Again using our example objects.

```{r, error=TRUE}
# a <- list(zahlen = 1:9, mon = month.abb, creator = "IQB")
# df <- data.frame(klein = 1:3, mittel = 4:6, groß = c("7", "8", "9"), row.names = c("A", "B", "C"))

assert_list(a, names = "named")
assert_data_frame(df, col.names = "unique")
```

You can check for specific names by using the `subset` functions. You can check whether the object give the argument `choices` contains `x`. If `x` is not a subset of `choices` you'll get an error message.

```{r, error=TRUE}
# lists
check_subset(x = "mon", choices = names(a))
# data frames
check_subset(c("klein", "mittel", "groß", "größer"), choices = colnames(df))

```

You can also check for unique values or missings in a similar way. For more info see the help functions.

```{r, eval=FALSE}
?assert_list
?assert_data_frame
```
